---
title: "Pstat 105 Lab2"
author: "Kendall Brown 8564403"
date: "Fall 2017"
output: pdf_document
---
Q1a. Table for all athletes in the data set.
Ho:Professional basketball players are born similarly to the general population.
Ha:Professional basketball players are born during a certain time of year.
Level=5%
```{r}
Bdays <- read.table("BBallBDays.txt")
Bday.month <- Bdays$Month
Bday.year <- Bdays$Year
Month.names=c("January","February","March","April","May","June","July",
"August","September","October","November","December")
Bday.month=factor(Bday.month,levels =Month.names,ordered = TRUE)
```

```{r}
mt=table(Bday.month)
mt
plot(mt,xlab="Month",ylab = "Number of Players")
```

Q1b.Chi-Squared test for all athletes. 
Ho: P(Jan)=P(Feb)=P(Mar)=...=P(Dec).
Ha: Each month has a different probability of birthing a professional basketball player.
```{r}
MEL=rep(1,12)/12
chisq.test(mt,p=MEL)
```
We observe a p-value of ~.26, indicating that there is not enough evidence to reject the null hypothesis at a 5% level.
There is not enough evidence to suggest that basketball players are not equaly likely to be born in each month.

Q2c.Table, Plot, and Chi-Squared test and table for athletes born after 1954.
Ho: For year>1954, P(Jan)=P(Feb)=P(Mar)=...=P(Dec).
Ha: Each month has a different probability of birthing a professional basketball player.
```{r}
Bday.1955=subset(Bdays,Year>1954)
Bday.month.1955=Bday.1955$Month
Bday.month.1955=factor(Bday.month.1955,levels = Month.names,ordered = TRUE)
mt1955=table(Bday.month.1955)
mt1955
plot(mt1955,xlab="Month",ylab="Number of Players Born After 1954")
chisq.test(mt1955,p=MEL)
```
We observe a p-value of ~.78, indicating that there is not enough evidence to reject the null hypothesis at a 5% level.
There is strong evidence to suggest that professional basketball players born after 1954 are equally likely to be born in each month  

Q1d. Chi-Squared Test considering days per month. 
Ho:Probabilty a professional basketball player is born in a certain monthis proportional to the number of days in that month.
Ha:Probability of a professional basketball player being born in a given month is independent of the number of days in that month.
```{r}
PPM=summary(Bday.month.1955)
DPM=c(31,28.25,31,30,31,30,31,31,30,31,30,31)
PDPM=DPM/365.25
chisq.test(PPM,p=PDPM)
```
We observe a p-value of ~.46, indicating that there is not enough evidence to reject the null hypothesis at a 5% level.
Professional basketball players appear to have a birthday distribution proportional to the average number of days in a given month over the average number of days in a given year.

Q1e. Chi-Squared test with CDC date.
Ho:Professional basketball players have birth months according to the CDC data.
Ha:Professional basketball players have birth months dissimilar to the CDC data.
```{r}
EPPMCDC=c(.0815,.0752,.0837,.0816,.0859,.0813,.0883,.0892,.0866,.0849,.0787,.0830)
chisq.test(x=PPM,p=EPPMCDC,rescale.p = TRUE)
```
We observe a p-value of ~.3, indicating that there is not enough evidence to reject the null hypothesis at a 5% level.
Professional basketball players appear to have a birthday distribution similar to the CDC estimates.

Q1f. From the tests conducted, we fail to reject the null hypothesis at a 5% level. Professional basketball players appear to share a similar birthday distribution to the general population.

Q2a. Histogram of Selltimes dataset.
```{r}
selltimes <- scan("Selltimes.txt")
hist.sell=hist((selltimes),breaks=35, main="Histogram of Selltimes",xlab = "Selltimes")
```

Q2b. MLE of Lambda
```{r}
lh=1/(mean(selltimes))
lh
```

Q2c. Histogram of Selltimes using MLE of Lambda.
```{r, message=FALSE, warning=FALSE}
#breakpoints chosen such that each box is expected to hold 10% of the data
bp=(log(seq(1,0,-.1)))/(-lh) 
bp
hist.sell.lh=hist(selltimes,breaks=bp,freq=T,xlim = c(0,250))
```

Q2d.Number of observations in each interval.
```{r}
hist.sell.lh$counts
```

Q2e. Chi-squared test.
```{r}
chisq.test(hist.sell.lh$counts)
```
Calculated p-value is near zero implying we reject the null hypothesis that the data follows an exponential distribution.

Q2f. Observed vs Expected counts
```{r}
hist.sell.lh$counts
chisq.test(hist.sell.lh$counts)$expected
```

It appears that, in comparison to our exponential, much of the observed data is skewed towards smaller values. Moreover, it is clearly shown that just under 2/3 of our observed values lay below our exponential distribution's lower 10% quantile.

Q2e. Testing various number of breaks
```{r, message=FALSE, warning=FALSE}
bp20=(log(seq(1,0,-.05)))/(-lh) #20 Breaks
hist.sell.lh.20=hist(selltimes,breaks=bp20,freq=T,xlim = c(0,250))
chisq.test(hist.sell.lh.20$counts)

bp40=(log(seq(1,0,-.025)))/(-lh) #40 Breaks
hist.sell.lh.40=hist(selltimes,breaks=bp40,freq=T,xlim = c(0,300))
chisq.test(hist.sell.lh.40$counts)

bp25=(log(seq(1,0,-.04)))/(-lh) #25 Breaks
hist.sell.lh.25=hist(selltimes,breaks=bp25,freq=T,xlim = c(0,250))
chisq.test(hist.sell.lh.25$counts)

bp50=(log(seq(1,0,-.02)))/(-lh) #50 Breaks
hist.sell.lh.50=hist(selltimes,breaks=bp50,freq=T,xlim = c(0,300))
chisq.test(hist.sell.lh.50$counts)
```
It appears as we increase the number of breaks in our histogram, the larger X-Squared test statistic we acquire. However, there is overwhelming evidence given here. We continue to reject the null hypothesis. The selltimes data set does not appear to follow an exponential distribution with parameter lambda = 0.01320614.





