---
title: "Pstat 150 Lab D"
author: "Kendall Brown"
date: "November 26, 2017"
output: pdf_document
---

```{r}
drugt=read.table("C:/Users/kebro/Desktop/PSTAT 105/Drug Treatment.txt",header=T)
library(survival)
```
Q1a.
Looking at the data in R, I see that after day 519 most of the observations became censored with only a few outliers being observed in the 600+ day range.

Q1b.
```{r}
dtc=Surv(drugt$Days,drugt$Censor)
dsf=survfit(dtc~1)
dsfs=summary(dsf)
dsfs$surv
plot(dsf,main="KM Estimator per Day",xlab="Days",ylab="KM Est.")
```

Q1b. 
```{r}
dsfecdf=1-dsfs$surv
dsqt=quantile(dsfs$time,c(.25,.5,.75))
dsqe=quantile(dsfs$surv,c(.25,.5,.75))
plot(dsfs$time,dsfecdf,type="s",xlab="Days",ylab="ECDF",col="red")
abline(h=dsqe,col=c("blue","yellow","green3"))
dsqe
dsqt
```
We have the .25, .5, and .75 quantiles of days to be 81, 167, and 276 respectively.

Q1d.
```{r}
drugpre=subset(drugt,IVDrug=="Previous",data=drugt)
drugrec=subset(drugt,IVDrug=="Recent",data=drugt)
drugnev=subset(drugt,IVDrug=="Never",data=drugt)
dtpc=Surv(drugpre$Days,drugpre$Censor)
dtrc=Surv(drugrec$Days,drugrec$Censor)
dtnc=Surv(drugnev$Days,drugnev$Censor)
dspf=survfit(dtpc~1)
dsrf=survfit(dtrc~1)
dsnf=survfit(dtnc~1)
dspfs=summary(dspf)
dsrfs=summary(dsrf)
dsnfs=summary(dsnf)
dspqt=quantile(dspfs$time,c(.25,.5,.75))
dsrqt=quantile(dsrfs$time,c(.25,.5,.75))
dsnqt=quantile(dsnfs$time,c(.25,.5,.75))
##.25,.5,.75 Quantiles for the previous, recent, and never observations in terms of time
dspqt
dsrqt
dsnqt
```

Q1e.
```{r}
dspqe=quantile(dspfs$surv,c(.25,.5,.75))
dsrqe=quantile(dsrfs$surv,c(.25,.5,.75))
dsnqe=quantile(dsnfs$surv,c(.25,.5,.75))
dspqe[1]
dsrqe[1]
dsrqe[1]
##95% Conf int for KM estimator for previous IV users after 75% mark
c(.295,.474)
##95% Conf int for KM estimator for recent IV users after 75% mark
c(.2508,.363)
##95% Conf int for KM estimator for people that have never had an IV after 75% mark
c(.255,.374)
```
From these confidence intervals, It can be said that there does not exist a significant differance between groups.
Q2.
```{r}
lungt=read.table("C:/Users/kebro/Desktop/PSTAT 105/lung.txt",header=T)
library(survival)
```
Q2a.
```{r}
lungsurv=Surv(lungt$time,abs(-1*lungt$status+1))
lsf=survfit(lungsurv~1)
lsfs=summary(lsf)
plot(lsf,main="KM Estimation Per Day",xlab = "Days",ylab="KM Est.")
```

Q2b.
```{r}
summary(lsf)
##95% conf int of KM estimator for 150 days
c(.7421,.848)
```

Q2c.
```{r}
##95% conf int of median survival time
lsf
```
we have a 95% conf int of {285,363}
Q2d.
```{r}
lungmale=subset(lungt,lungt$sex==1)
lungfemale=subset(lungt,lungt$sex==2)
lungmalesf=Surv(lungmale$time,abs(-1*lungmale$status+1))
lungfemalesf=Surv(lungfemale$time,abs(-1*lungfemale$status+1))
lungmfs=survfit(lungmalesf~1)
lungffs=survfit(lungfemalesf~1)
plot(lungffs,main="Female KM Estimation Per Day",xlab = "Days",ylab="KM Est.")
plot(lungmfs,main="Male KM Estimation Per Day",xlab = "Days",ylab="KM Est.")
```
From these plot it appears that women have higher survival rates seemingly along the entire

Q2e. MvF median surv times
```{r}
lungmfs
lungffs
```
From these confidence intervals and medians, I believe it is rather clear that women tend to survive longer than men. However, i do not believe that this tells the whole story as other factors not measured in this data set could be rather impactful to survival time.
Q3.1,000,000 sims
```{r}
simexp=rexp(10^6)
cossim=cos(simexp)
cossim.7=(cossim>.7)/1
cossim.7h=hist(cossim.7,breaks=c(0,.5,1))
cossim.7p=cossim.7h$counts[2]/(10^6)
cossim.7p ##Probability Cos(x)>.7
##95% conf int for P(cos(x)>.7)
c(cossim.7p-1.96*sqrt(cossim.7p*(1-cossim.7p)/(10^6)),
  cossim.7p+1.96*sqrt(cossim.7p*(1-cossim.7p)/(10^6)))
```

Q4a.1000 runs of 50 sims
```{r}
n=50
m=1000
ex.test=rep(999,m)
rnd.x=matrix(rep(999,50000),1000,50)
for(j in 1:m){
  rnd.x[j,]=c(rexp(n))
  ex.test[j]=mean(rnd.x[j,])
}
hist(ex.test,breaks=30,col="red2",main="Hist from Simulation",xlab="Calculated Means")
```

Q4b. KS-test of samples
```{r}
kssim=rep(999,1000)
for(j in 1:m){
  ksdj=ks.test(rnd.x[j,],pexp)
  kssim[j]=ksdj$statistic
}
histks=hist(kssim,breaks=30,main="Simulated KS-test Statistics",xlab="D Statistics",col="blue")
```

Q4c.
```{r}
quantile(kssim,.95)
```

Q4d.
```{r}
#computer ran a single batch in .16 seconds 
#7200 secs in 2hrs means 45000 batches
#to avoid time issues when knitting this block will be ## out

#kssim2h=rep(999,m)
#rnd2h.x=matrix(rep(999,50000),1000,50)
#batch=matrix(rep(999,4.5e+07),45000,1000)
#for(j in 1:45000){
#for(k in 1:m){
  #rnd2h.x[k,]=c(rexp(n))
  #ex.test[k]=mean(rnd2h.x[k,])
  #ksdj2h=ks.test(rnd2h.x[k,],pexp)
  #kssim2h[k]=ksdj2h$statistic
  #batch[j,]=kssim2h
#}}
```

```{r}
#sim2h=quantile(batch,.95)
#sim2h
```
To avoid problems during knitting, I am stating the calculated simulation .95 quantile to be .1884515.

Q4e.Crit value from Stephens table 1.4
```{r}
(.1884515-.2/50)*(sqrt(50)+.26+.5/sqrt(50))
```
From this calculation we get a T value that corresponds to the <1% level.  

Q5a.
```{r}
library(MASS)
ofsd=sd(geyser$waiting)
ofsd ##sample standard deviation of geyser data set
```

Q5b.
```{r}
geysermat=rep(999,1000)
for(j in 1:1000){
sampof=sample(geyser$waiting,299,replace = T)
geysermat[j]=sd(sampof)
}
(sd(geysermat))^2
```

Q5c.
```{r}
geyseriqr=rep(999,1000)
for(j in 1:1000){
sampofiqr=sample(geyser$waiting,299,replace = T)
geyseriqr[j]=IQR(sampofiqr)
}
(sd(geyseriqr))^2
```

Q5d.
```{r}
geysermad=rep(999,1000)
for(j in 1:1000){
sampofmad=sample(geyser$waiting,299,replace = T)
geysermad[j]=mad(sampofmad)
}
(sd(geysermad))^2
```

```{r}
hist(geysermat,breaks=100)
hist(geyseriqr,breaks=100)
hist(geysermad,breaks=100)
```
Based on the histograms of the standard deviation estimates, the SD function has the most normaly distributed data. As such, personally, i prefer the sd function over the other two estimation functions.