---
title: "Pstat 131 Homework 3"
author: "Kendall Brown 8564403"
date: "Winter 2018"
output: pdf_document
---
```{r}
library(tidyverse)
library(e1071)
library(cluster)
library(NbClust)
library(bmp)
library(imager)
library(tree)
library(randomForest)
library(gbm)
library(ROCR)
```
1a. In a sample of size n, the probability that the jth observation is not in the bootstrap is the product of
the probability that each observation in the bootstrap is not the jth element of the sample. If each element
has probability 1
n of being in the bootstrap, then each element has probability $(1 ??? \frac{1}{n})$ of not being in the
bootstap. Thus the probability that the jth observation is not in the bootstrap is
$(1 ??? \frac{1}{n})^n$

1b. Probability when n=1000
```{r}
(1-1/1000)^1000
```

1c. Resampling numbers from 1 to 1000, then calculating observed probability that an element is not part of
the bootstrap sample.
```{r}
set.seed(300)
n.1c=c(1:1000)
sample.1c=sample(n.1c,1000,replace = T)
sample.1c.unique=unique(sample.1c)
p.n.1c=length(sample.1c.unique)/1000
1-p.n.1c
```

Setting the seed to be 300, I observe that the probability of an element not being in the bootstrap is .366.

1d. Basketball shot field goal precentage.
```{r}
n.1d=c(rep(1:50),rep(0:51))
fgp.1d=rep(0,1000)
for(i in 1:1000){
sample.1d=sample(n.1d,101,replace=T)
fgp.1d[i]=mean(sample.1d)/101
}
```

Histogram of Field Goal Percentages from a bootstrap sample of 101 shots. Overlayed with normality curve
and 95% confidence interval.
```{r}
fgp.ci=quantile(fgp.1d,probs=c(.025,.975))
hist(fgp.1d,breaks=15,col="red",
main="FGPs over 101 Shots with Normality curve and 95% C.I.",
xlab="Field Goal Percentage",probability = T)
curve(dnorm(x,mean(fgp.1d),sd(fgp.1d)),add=T,col="blue")
abline(v=fgp.ci,col="blue")
```
From the histogram and overlayed gausian curve, we see a strong indication of normality with little skew.
95% confidence interval for Rober Covington's FGP

```{r}
fgp.ci#95% confidence Interval
```

```{r}
fgp.ci.9999=quantile(fgp.1d,probs=c(.9999))#Upper 99.99% confidence bound of FGPs
fgp.ci.9999
```

From this 95% confidence interval I must assume that Robert Covington's FGP will fall by the end of the
competitive season. As of now he is shooting well beyond the realm of expectation with an average of .495.
Such a value dwarfs the 99.99% confidence bound of .299 and should be viewed as a statistical outlier.

```{r}
seeds = read.table(
  'http://archive.ics.uci.edu/ml/machine-learning-databases/00236/seeds_dataset.txt')
names(seeds) = c('area', 'perimeter', 'compactness', 'length',
'width', 'asymmetry', 'groovelength', 'type')
seeds.label = factor(seeds$type)
seeds.orig = seeds[,-ncol(seeds)]
seeds = scale(seeds.orig)
seeds = as.data.frame(seeds)
```
