---
title: "Untitled"
output: html_document
---

```{r}
rm(list=ls())
library(readxl)
library(tidyverse)
library(hash)
library(lubridate)
library(caret)
library(e1071)
library(PRROC)
```

```{r}
# reservations=read_xlsx("C:/Users/kebro/OneDrive/KU_Leuven/StatCons/reservations.xlsx")
reservations=as_tibble(read.csv("C:/Users/kebro/OneDrive/KU_Leuven/StatCons/reservations.csv",header=T))
```
```{r}
str(reservations)
```

clean arrival date variable
```{r}
datevars=reservations%>%
  select(arrival_date_day_of_month,arrival_date_month,arrival_date_year)
```

```{r}
monthnumber=hash()
months=c("January","February","March","April","May","June",
         "July","August","September","October","November","December")
for(i in 1:12){
  monthnumber[months[i]]=i
}
```

```{r}
datevars$monthnum=rep(0,length(datevars$arrival_date_month))
for(i in 1:length(datevars$arrival_date_month)){
  datevars$monthnum[i]=monthnumber[[as.character(datevars$arrival_date_month[i])]]
}
```

```{r}
arrivaldate=rep("",length(datevars$arrival_date_month))
for(i in 1:length(arrivaldate)){
  arrivaldate[i]=paste(datevars[i,1],datevars[i,4],datevars[i,3],sep="-")
}
```
```{r}
datevars=datevars%>%mutate(arrival=dmy(arrivaldate))
reservations=reservations%>%mutate(arrival_date=datevars$arrival)
```

```{r}
dates=unique(reservations$arrival_date)
x=length(dates)
reservations_per_day=rep(0,x)
cancelations_per_day=rep(0,x)
for(i in 1:x){
  reservations_per_day[i]=length(reservations$arrival_date[reservations$arrival_date==dates[i]])
  cancelations_per_day[i]=sum(reservations$is_canceled[reservations$arrival_date==dates[i]])
}
```

```{r}
per_day_statistics=tibble(
  dates,reservations_per_day,
  cancelations_per_day,
  cancelations_per_day_ratio=cancelations_per_day/reservations_per_day)
```

```{r}
per_day_statistics%>%
  ggplot(aes(dates,cancelations_per_day_ratio))+
  geom_point()+
  labs(y="Percent Cancelations",x="Date",title="Global")
```

```{r}
x="USA"
reservations_x=reservations%>%filter(country==x)
dates_x=unique(reservations_x$arrival_date)
x=length(dates_x)
reservations_per_day=rep(0,x)
cancelations_per_day=rep(0,x)
for(i in 1:x){
  reservations_per_day[i]=length(reservations_x$arrival_date[reservations_x$arrival_date==dates_x[i]])
  cancelations_per_day[i]=sum(reservations_x$is_canceled[reservations_x$arrival_date==dates_x[i]])
}
per_day_statistics=tibble(
  dates_x,reservations_per_day,
  cancelations_per_day,
  cancelations_per_day_ratio=cancelations_per_day/reservations_per_day)
per_day_statistics%>%
  ggplot(aes(dates_x,cancelations_per_day_ratio))+
  geom_point()+
  labs(y="Percent Cancelations",x="Date",title="USA")
```

```{r}
x="BEL"
reservations_x=reservations%>%filter(country==x)
dates_x=unique(reservations_x$arrival_date)
x=length(dates_x)
reservations_per_day=rep(0,x)
cancelations_per_day=rep(0,x)
for(i in 1:x){
  reservations_per_day[i]=length(reservations_x$arrival_date[reservations_x$arrival_date==dates_x[i]])
  cancelations_per_day[i]=sum(reservations_x$is_canceled[reservations_x$arrival_date==dates_x[i]])
}
per_day_statistics=tibble(
  dates_x,reservations_per_day,
  cancelations_per_day,
  cancelations_per_day_ratio=cancelations_per_day/reservations_per_day)
per_day_statistics%>%
  ggplot(aes(dates_x,cancelations_per_day_ratio))+
  geom_point()+
  labs(y="Percent Cancelations",x="Date",title = "Belgium") 
```
```{r}
x="CHN"
reservations_x=reservations%>%filter(country==x)
dates_x=unique(reservations_x$arrival_date)
x=length(dates_x)
reservations_per_day=rep(0,x)
cancelations_per_day=rep(0,x)
for(i in 1:x){
  reservations_per_day[i]=length(reservations_x$arrival_date[reservations_x$arrival_date==dates_x[i]])
  cancelations_per_day[i]=sum(reservations_x$is_canceled[reservations_x$arrival_date==dates_x[i]])
}
per_day_statistics=tibble(
  dates_x,reservations_per_day,
  cancelations_per_day,
  cancelations_per_day_ratio=cancelations_per_day/reservations_per_day)
per_day_statistics%>%
  ggplot(aes(dates_x,cancelations_per_day_ratio))+
  geom_point()+
  labs(y="Percent Cancelations",x="Date",title="China")
```
```{r}
x="PRT"
reservations_x=reservations%>%filter(country==x)
dates_x=unique(reservations_x$arrival_date)
x=length(dates_x)
reservations_per_day=rep(0,x)
cancelations_per_day=rep(0,x)
for(i in 1:x){
  reservations_per_day[i]=length(reservations_x$arrival_date[reservations_x$arrival_date==dates_x[i]])
  cancelations_per_day[i]=sum(reservations_x$is_canceled[reservations_x$arrival_date==dates_x[i]])
}
per_day_statistics=tibble(
  dates_x,reservations_per_day,
  cancelations_per_day,
  cancelations_per_day_ratio=cancelations_per_day/reservations_per_day)
per_day_statistics%>%
  ggplot(aes(dates_x,cancelations_per_day_ratio))+
  geom_point()+
  labs(y="Percent Cancelations",x="Date",title="Portugal")
```

```{r}
x="FRA"
reservations_x=reservations%>%filter(country==x)
dates_x=unique(reservations_x$arrival_date)
x=length(dates_x)
reservations_per_day=rep(0,x)
cancelations_per_day=rep(0,x)
for(i in 1:x){
  reservations_per_day[i]=length(reservations_x$arrival_date[reservations_x$arrival_date==dates_x[i]])
  cancelations_per_day[i]=sum(reservations_x$is_canceled[reservations_x$arrival_date==dates_x[i]])
}
per_day_statistics=tibble(
  dates_x,reservations_per_day,
  cancelations_per_day,
  cancelations_per_day_ratio=cancelations_per_day/reservations_per_day)
per_day_statistics%>%
  ggplot(aes(dates_x,cancelations_per_day_ratio))+
  geom_point()+
  labs(y="Percent Cancelations",x="Date")
```

```{r}
x="ITA"
reservations_x=reservations%>%filter(country==x)
dates_x=unique(reservations_x$arrival_date)
x=length(dates_x)
reservations_per_day=rep(0,x)
cancelations_per_day=rep(0,x)
for(i in 1:x){
  reservations_per_day[i]=length(reservations_x$arrival_date[reservations_x$arrival_date==dates_x[i]])
  cancelations_per_day[i]=sum(reservations_x$is_canceled[reservations_x$arrival_date==dates_x[i]])
}
per_day_statistics=tibble(
  dates_x,reservations_per_day,
  cancelations_per_day,
  cancelations_per_day_ratio=cancelations_per_day/reservations_per_day)
per_day_statistics%>%
  ggplot(aes(dates_x,cancelations_per_day_ratio))+
  geom_point()+
  labs(y="Percent Cancelations",x="Date",title="Italy")
```

```{r}
x="DEU"
reservations_x=reservations%>%filter(country==x)
dates_x=unique(reservations_x$arrival_date)
x=length(dates_x)
reservations_per_day=rep(0,x)
cancelations_per_day=rep(0,x)
for(i in 1:x){
  reservations_per_day[i]=length(reservations_x$arrival_date[reservations_x$arrival_date==dates_x[i]])
  cancelations_per_day[i]=sum(reservations_x$is_canceled[reservations_x$arrival_date==dates_x[i]])
}
per_day_statistics=tibble(
  dates_x,reservations_per_day,
  cancelations_per_day,
  cancelations_per_day_ratio=cancelations_per_day/reservations_per_day)
per_day_statistics%>%
  ggplot(aes(dates_x,cancelations_per_day_ratio))+
  geom_point()+
  labs(y="Percent Cancelations",x="Date",title="Germany")
```
```{r}
x="SWE"
reservations_x=reservations%>%filter(country==x)
dates_x=unique(reservations_x$arrival_date)
x=length(dates_x)
reservations_per_day=rep(0,x)
cancelations_per_day=rep(0,x)
for(i in 1:x){
  reservations_per_day[i]=length(reservations_x$arrival_date[reservations_x$arrival_date==dates_x[i]])
  cancelations_per_day[i]=sum(reservations_x$is_canceled[reservations_x$arrival_date==dates_x[i]])
}
per_day_statistics=tibble(
  dates_x,reservations_per_day,
  cancelations_per_day,
  cancelations_per_day_ratio=cancelations_per_day/reservations_per_day)
per_day_statistics%>%
  ggplot(aes(dates_x,cancelations_per_day_ratio))+
  geom_point()+
  labs(y="Percent Cancelations",x="Date")
```

Baseline Model: this is proving to be unfruiful as even after significant backward elimination the model suffers from complete seperation and a slew of other bugs best to to move towards a forward selction method in a naive model.
```{r}
set.seed(123)
reservations=reservations%>%mutate(id=row_number())
reservationbase=reservations%>%filter(arrival_date>as.Date("2016-01-01"))

reservationbase=reservationbase%>%
  select(-reservation_status,-reservation_status_date,
         -country,-arrival_date_year,-arrival_date_month,
         -arrival_date_day_of_month,-agent,-company,
         -reserved_room_type,-assigned_room_type,-arrival_date_week_number,
         -required_car_parking_spaces,-booking_changes,
         -market_segment,-lead_time,-previous_cancellations,-deposit_type,
         -total_of_special_requests,-hotel,-customer_type,-arrival_date)

trainbase=reservationbase%>%sample_frac(.8)
testbase=anti_join(reservationbase,trainbase,by="id")
trainbase=trainbase%>%select(-id)
testbase=testbase%>%select(-id)
```



```{r}
base_model=glm(is_canceled~.,trainbase,family = "binomial")
summary(base_model)
```

```{r}
test_resultsbase=predict.glm(base_model,newdata = testbase,type="response")
class_imb=sum(as.integer(trainbase$is_canceled)-1)/length(trainbase$is_canceled)
decision_bound=class_imb
predicted_test_classesbase=rep(0,length(test_resultsbase))
for(i in 1:length(predicted_test_classesbase)){
  if(test_resultsbase[i]<=1-decision_bound){
    predicted_test_classesbase[i]=0
  }
  if(test_resultsbase[i]>decision_bound){
    predicted_test_classesbase[i]=1
  }
}
predicted_test_classesbase=as_factor(predicted_test_classesbase)
confusionMatrix(predicted_test_classesbase,as_factor(testbase$is_canceled))
```

Making a naive model based off of forward selection in hopes of avoiding complete seperation.
```{r}
set.seed(123)
updated_reservations=tibble(.rows = nrow(reservations))
updated_reservations=updated_reservations%>%
  mutate(is_canceled=as.factor(reservations$is_canceled))%>%
  mutate(adr=(reservations$adr-mean(reservations$adr))/sd(reservations$adr))%>%
  mutate(lead_time=(reservations$lead_time-mean(reservations$lead_time))/sd(reservations$lead_time))%>%
  mutate(hotel=reservations$hotel)%>%
  mutate(id=row_number())%>%
  mutate(arrival_date=reservations$arrival_date)%>%
  filter(arrival_date>as.Date("2017-03-01"))
train0=updated_reservations%>%sample_frac(.8)
test0=anti_join(updated_reservations,train0,by="id")
train0=train0%>%select(-id)
test0=test0%>%select(-id)
```

```{r}
naive_model=glm(is_canceled~.,data=train0,family = "binomial")
summary(naive_model)
```


```{r}
test_results0=predict.glm(naive_model,newdata = test0,type="response")
```

```{r}
class_imb=sum(as.integer(train0$is_canceled)-1)/length(train0$is_canceled)
```

```{r}
decision_bound=1-class_imb
predicted_test_classes0=rep(0,length(test_results0))
for(i in 1:length(predicted_test_classes0)){
  if(test_results0[i]<=1-decision_bound){
    predicted_test_classes0[i]=0
  }
  if(test_results0[i]>decision_bound){
    predicted_test_classes0[i]=1
  }
}
predicted_test_classes0=as_factor(predicted_test_classes0)
confusionMatrix(predicted_test_classes0,test0$is_canceled)
```
The complete seperation problem appears to have been rectified in this naive model, however the model evaluation metrics appear to be somewhat lack luster. Prehaps a somewhat larger model will give better results whilse still avoiding complete speration.


Making a bigger model
```{r}
reservations=reservations%>%
  mutate(total_guests=adults+children+babies)%>%
  mutate(child_check=as_factor(case_when(children>0|babies>0~1,
                                         children==0&babies==0~0)))%>%
  mutate(group_factor=as_factor(case_when(total_guests==0~"single",
                                total_guests==1~"single",
                                total_guests==2~"double",
                                total_guests>=3&total_guests<=5~"triple+",
                                total_guests>5~"large")))

price_category=tibble(.rows = nrow(reservations))
price_category=price_category%>%
  mutate(adr=reservations$adr)%>%
  mutate(category=as_factor(case_when(adr<=50~"economy",
                                 adr>50&adr<=150~"standard",
                                 adr>100&adr<=300~"expensive",
                                 adr>300~"luxury")))

room_diff=(as.character(reservations$assigned_room_type)==as.character(reservations$reserved_room_type))

set.seed(123)
updated_reservations=tibble(.rows = nrow(reservations))
updated_reservations=updated_reservations%>%
  mutate(is_canceled=as.factor(reservations$is_canceled))%>%
  mutate(adr=reservations$adr)%>%
  #mutate(adr=price_category$category)%>%
  #mutate(adr=(reservations$adr-mean(reservations$adr))/sd(reservations$adr))%>%
  mutate(lead_time=reservations$lead_time)%>%
  mutate(hotel=reservations$hotel)%>%
  #mutate(total_guests=reservations$total_guests)%>%
  mutate(group_factor=reservations$group_factor)%>%
  mutate(child_check=reservations$child_check)%>%
  mutate(is_repeated_guest=reservations$is_repeated_guest)%>%
  mutate(deposit_type=reservations$deposit_type)%>%
  mutate(meal=reservations$meal)%>%
  mutate(room_check=as_factor(room_diff))%>%
  mutate(distribution_channel=reservations$distribution_channel)%>%
  mutate(arrival_date_week_number=reservations$arrival_date_week_number)%>%
  mutate(country=reservations$country)%>%
  mutate(arrival_date=reservations$arrival_date)%>%
  filter(arrival_date>as.Date("2017-03-01"))%>%
  mutate(id=row_number())%>%
  group_by(country)%>%
  mutate(count=n())%>%
  filter(count>20)%>%
  select(-count)%>%
  droplevels()
train1=updated_reservations%>%sample_frac(.8)
test1=anti_join(updated_reservations,train1,by="id")
train1=train1%>%select(-id)
test1=test1%>%select(-id)
```

```{r}
bigger_model=glm(is_canceled~.+adr*room_check+child_check*adr+arrival_date*country,data=train1,family = "binomial")
summary(bigger_model)
```

```{r}
highdev=(residuals.glm(bigger_model,type="deviance")<2)&(-2<residuals.glm(bigger_model,type="deviance"))
train_nohighdev=train1[highdev,]
```

```{r}
bigger_model=glm(is_canceled~.+adr*room_check+child_check*adr+arrival_date*country,data=train_nohighdev,
                 family = "binomial")
summary(bigger_model)
```




```{r}
test_results1=predict.glm(bigger_model,newdata = test1,type="response")
plot(roc.curve(scores.class0 = 1-test_results1,weights.class0 = (as.numeric(test1$is_canceled)-2)*-1,curve=T))
```


```{r}
logs=rep(0,length(seq(0,1,by=.01)))
count=1
for(j in seq(0,1,by=.01)){
  decision_bound=j
  predicted_test_classes1=rep(0,length(test_results1))
  for(i in 1:length(predicted_test_classes1)){
    if(test_results1[i]<=1-decision_bound){
      predicted_test_classes0[i]=0
    }
    if(test_results1[i]>decision_bound){
      predicted_test_classes1[i]=1
    }
  }
  predicted_test_classes1=as_factor(predicted_test_classes1)
  levels(predicted_test_classes1)=c(0,1)
  x=confusionMatrix(predicted_test_classes1,test1$is_canceled)
  logs[count]=x$byClass[length(x$byClass)][[1]]
  count=count+1
}
plot(y=logs,x=seq(0,1,by=.01))
```

```{r}
decision_bound=.33
predicted_test_classes1=rep(0,length(test_results1))
for(i in 1:length(predicted_test_classes1)){
  if(test_results1[i]<=1-decision_bound){
    predicted_test_classes0[i]=0
  }
  if(test_results1[i]>decision_bound){
    predicted_test_classes1[i]=1
  }
}
predicted_test_classes1=as_factor(predicted_test_classes1)
confusionMatrix(predicted_test_classes1,test1$is_canceled)
```
```{r}
exp(coef(bigger_model))
```
```{r}
reservations%>%
  ggplot(aes(x=deposit_type,fill=factor(is_canceled)))+
  geom_bar(position="fill")+
  labs(x="Deposit Type",y="Proportion of Reservations")+
  scale_fill_discrete(name="Cancellation",labels=c("No","Yes"))
```
```{r}
updated_reservations%>%
  ggplot(aes(x=is_repeated_guest,fill=factor(is_canceled)))+
  geom_bar(position="fill")+
  labs(x="Repeated Guest",y="Proportion of Reservations")+
  scale_fill_discrete(name="Cancellation",labels=c("No","Yes"))
```
```{r}

```