---
title: "Seminar 1 - GLM"
author: "Group #01"
date: "03/17/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

A data frame with 75 observations on the following 4 variables:

group: A factor with 2 levels CTRL and TREAT, indicating the treatment.
n: A numeric vector indicating the number of animals exposed to mortality.
trisk: A numeric vector indicating the exposition time to mortality (in year). This variable should not be used in the analysis.
y: A numeric vector indicating the number of deaths.
Scientific question: Does treatment (group) have an effect on mortality?

1.Fit a logistic regression model.
2.Calculate the odds ratios and corresponding confidence intervals.
3.Assess the GOF of the model (Pearson and Deviance test).
4.Predicted probability of death for each treatment.
5.Fit a quasilikelihood model. Comment.
6.Fit a beta binomial model. Comment.
```{r}
rm(list=ls())
library(dplyr)
```

```{r}
djasem <- read.table("C:/Users/kebro/OneDrive/KU_Leuven/GLM/djaSem.txt", header = T)
attach(djasem)
data<-na.omit(subset(djasem,select=c("group","n", "y")))
data$group1<-as.numeric(as.factor(djasem$group))-1
sum(data$group1)
as.factor(data$group1)
data$prop<-y/n
detach(djasem)
attach(data)
```
descriptive statistics
```{r exploration, echo=FALSE}
cor(data[ ,-1])
summary(data)
```
Logistics model
```{r logistics model,echo=FALSE}
treat.logit.1<-glm(cbind(y, n-y)~group1, family=binomial(link="logit"),data=data)
summary(treat.logit.1)
```
Grapical display
```{r logistic curve,echo=FALSE,warning=FALSE}
plot(group1,prop,ylim=c(0,1.05),type="n",
cex=1., xlab="group",ylab="Proportion of Death")
points(group1,prop,pch=19,col="red")
## Plotting the logistic curve
curve(predict(treat.logit.1,data.frame(group1=x),type="resp"),
add=TRUE,col="steelblue",lwd=3)
```
odds ratio and interval
```{r odds ratio, echo=FALSE,warning=FALSE}
oddsratio<-exp(treat.logit.1$coefficients)
limit<-exp(confint(treat.logit.1))
interval<-exp(cbind(OR=treat.logit.1$coefficients, confint(treat.logit.1)))
interval
```
Pearon and Deviance test
```{r pearson test,echo=FALSE,warning=FALSE}
X2=sum(residuals(treat.logit.1, type = "pearson")^2)
data.frame(X2s=X2,pvalue=(1-pchisq(X2,73)))

attributes(summary(treat.logit.1))
Dev.an=summary(treat.logit.1)$deviance
df.an=summary(treat.logit.1)$df.residual
data.frame(Dev=Dev.an, df=df.an, pvalue=(1-pchisq(Dev.an,df.an)))
```
Both pearson Chi-square test and Deviance test reject the model.
We will check the residual plot.
Residual analysis
```{r residual plot,echo=FALSE,warning=FALSE}
treat.rdev<-residuals(treat.logit.1,type="deviance")
 plot(predict(treat.logit.1),treat.rdev,xlab="Linear predictor", ylab="Devianceresiduals")
abline(h=0,lty=2,col="grey")
loess.treat.logit.1.rdev<-loess(treat.rdev~predict(treat.logit.1))
treat.logit.1.lo.pred.dev <- predict(loess.treat.logit.1.rdev, se=T)
j.treat<- order(predict(treat.logit.1))
lines(predict(treat.logit.1)[j.treat],treat.logit.1.lo.pred.dev$fit[j.treat],col="blue",lwd=3)
lines(predict(treat.logit.1)[j.treat],treat.logit.1.lo.pred.dev$fit[j.treat]+2*treat.logit.1.lo.pred.dev$s[j.treat], lty=2,col="red")
lines(predict(treat.logit.1)[j.treat],treat.logit.1.lo.pred.dev$fit[j.treat]-2*treat.logit.1.lo.pred.dev$s[j.treat], lty=2,col="red")
```
residual plot :group1
```{r group1, echo=FALSE,warning=FALSE}

treat.logit.1rdev<-residuals(treat.logit.1,type="deviance")
plot(data$group1,treat.logit.1rdev,xlab="Group1", ylab="Deviance residuals")
abline(h=0,lty=2,col="grey")
loess.treat.logit.1rdev.group1 <- loess(treat.logit.1rdev~data$group1)
treat.logit.1lo.pred.dev.group1 <- predict(loess.treat.logit.1rdev.group1, se=T)
 j.treat.group1<- order(data$group1)
 lines(data$group1[j.treat.group1],treat.logit.1lo.pred.dev.group1$fit[j.treat.group1],
col="blue",lwd=3)
 lines(data$group1[j.treat.group1],treat.logit.1lo.pred.dev.group1$fit[j.treat.group1]
+2*treat.logit.1lo.pred.dev.group1$s[j.treat.group1], lty=2,col="red")
 lines(data$group1[j.treat.group1],treat.logit.1lo.pred.dev.group1$fit[j.treat.group1]
 -2*treat.logit.1lo.pred.dev.group1$s[j.treat.group1], lty=2,col="red")
```
influential points 
```{r outliers, echo=FALSE}
treat.logit.1.hat<-hatvalues(treat.logit.1)
treat.logit.1.rs <- rstudent(treat.logit.1)
treat.logit.1.st <- residuals(treat.logit.1, type="pearson")/sqrt(1 - treat.logit.1.hat)
deltaX2 <- treat.logit.1.st^2
deltaD <- treat.logit.1.hat*treat.logit.1.st^2+residuals(treat.logit.1, type="deviance")^2
plot(deltaX2,xlab="Case number",ylab="Delta Chi-Square", col='red',pch=20)
lines(deltaX2)
identify(deltaX2)
plot(deltaD,xlab="Case number",ylab="Delta Deviance", col='red',pch=20)
lines(deltaD)
identify(deltaD)

```
prediction based on Logistics regression
```{r prediction, echo=FALSE}
new<-data.frame(group1=c(0,1))
predict(treat.logit.1,newdata = new,type = 'response')
```
Biased reduction LR - Quasi-likelihood model
```{r quasi,echo=FALSE}
library(brglm)
treat.logit.2<-brglm(cbind(y, n-y)~group1, family=binomial(link="logit"),data=data)
summary(treat.logit.2)
```
Test for BRLR
```{r test2,echo=FALSE}
X2=sum(residuals(treat.logit.2, type = "pearson")^2)
data.frame(X2s=X2,pvalue=(1-pchisq(X2,73)))
attributes(summary(treat.logit.2))
Dev.an=summary(treat.logit.2)$deviance
df.an=summary(treat.logit.2)$df.residual
data.frame(Dev=Dev.an, df=df.an, pvalue=(1-pchisq(Dev.an,df.an)))
```
Quality of Prediction 
```{r r^2,echo=FALSE}
library(DescTools)
require(BaylorEdPsych)
PseudoR2(treat.logit.1)
cbind(LR=PseudoR2(treat.logit.1)[4],BRLR=PseudoR2(treat.logit.2)[4])
cbind(LR=PseudoR2(treat.logit.1)[3],BRLR=PseudoR2(treat.logit.2)[3])
```
Test for overdispersion
```{r overdispersion}
library(DHARMa)
sim.model.ab<- simulateResiduals(treat.logit.1, refit=T)
testDispersion(sim.model.ab)
```
beta binomial 
```{r beta, echo=FALSE}
library(aod)# aod betabin
library(car)
treat.beta.1<-betabin(cbind(y, n-y)~group1,~1,link = "logit",data=data )

View(data)
str(data)
summary(treat.beta.1)
vcov(treat.beta.1)
logLik(treat.beta.1)
AIC(treat.beta.1)
coef(treat.beta.1)
```
test for beta binomial
lower.trail = F gives the same result.
```{r GOF test,echo=FALSE}
X2.ab.nb=sum(residuals(treat.beta.1, type = "pearson")^2)
n.ab.nb=dim(data)[1]
p.ab.nb=length(coef(treat.beta.1))
data.frame(X2s=X2.ab.nb,pvalue=1-pchisq(X2.ab.nb,n.ab.nb-p.ab.nb))
```
beta-binomial prediction
```{r prediction, echo=FALSE}
new<-data.frame(group1=c(0,1))
predict(treat.beta.1,newdata = new,type = 'response')
```

Average Mortality rate from dataset 
```{r Mortality Rate, echo=FALSE}

#Population Weighted Mortality rate
library(dplyr)

Only.Treat<-filter(data,group1==1)
nrow(Only.Treat)
Treat.rate<-sum(Only.Treat$y)/sum(Only.Treat$n)

Only.Ctrl<-filter(dataset,group1==0)
nrow(Only.Ctrl)
Ctrl.rate<-sum(Only.Ctrl$y)/sum(Only.Ctrl$n)

data.frame(Ctrl.rate,Treat.rate)

####################################

# Unweighted Mortality rate

data2=dataset%>%mutate(prop=y/n)
mean(data2$prop[group=="TREAT"])
mean(data2$prop[group=="CTRL"])
```